---
title: "L2"
author: "COMPASS workflows team"
title-block-banner: true
params:
  html_outfile: "L2.html"
  DATA_ROOT: "data_TEST/"
  L2: "L2/"
  L2_QAQC: "L2_qaqc/"
  QAQC_TABLE: "qaqc_table.csv"
  timestamp_round: "15 min"
  METADATA_ROOT: "metadata/"
  L2_METADATA: "L2_metadata/"
  L2_RELEASE_DATE: "???"
  L2_DATA_TIMEZONE: "Etc/GMT+5"
  L2_VERSION: "???"
  run_parallel: false
  write_plots: false
  logfile: ""
date: now
date-format: "YYYY-MM-DD HH:mm:ssZ"
format: 
  html:
    embed-resources: true
    code-fold: true
editor: visual
---

This script

-   Reads in the L2_qaqc data files one by one

-   Computes derived variables

-   Calls gap-filling algorithms (currently only mean annual cycle)

-   Writes L2 files and accompanying documentation

## Initializing

```{r init}
#| include: false

library(tidyr)
library(readr)
library(lubridate)
library(zoo)
library(ggplot2)


L2_QAQC <- file.path(params$DATA_ROOT, params$L2_QAQC)
L2 <- file.path(params$DATA_ROOT, params$L2)

files_to_process <- list.files(L2_QAQC, 
                               pattern = "[A-Z]{3}_[A-Z]{1,2}_[0-9]{8}-[0-9]{8}_.+csv$", 
                               full.names = TRUE, recursive = TRUE)

QAQC_TABLE <- file.path(params$METADATA_ROOT, params$QAQC_TABLE)
qaqct <- read_csv(QAQC_TABLE, col_types = "cccccc")

source("helpers.R")
source("L2_qaqc-utils.R")
```

I see `r length(files_to_process)` files to process in `r L2_QAQC`.

QAQC table `r QAQC_TABLE` and has `r nrow(qaqct)` entries.

Output directory is `r L2`

HTML outfile is "`r params$html_outfile`".

## Processing

```{r processing}

# something
errors <- 0
out <- head(cars)
```

## Summary

```{r summary}
#| echo: false
#| output: asis
if(errors) {
    cat("### WARNING: ", errors, " file read/write error(s)\n")
    log_warning(paste("File read/write error(s)", params$html_outfile), 
                logfile = params$logfile)
}
```

```{r summary_table}
out_df <- do.call("rbind", out)
knitr::kable(out_df[-1]) # don't show filename
```

## Reproducibility

Git commit `r GIT_COMMIT`.

```{r reproducibility}
sessionInfo()
```
