---
title: "L2_qaqc"
author: "COMPASS workflows team"
title-block-banner: true
params:
  html_outfile: "L2_qaqc.html"
  DATA_ROOT: "data_TEST/"
  L1: "L1/"
  L2_QAQC: "L2_qaqc/"
  QAQC_TABLE: "qaqc_table.csv"
  timestamp_round: "15 min"
  METADATA_ROOT: "metadata/"
  L2_METADATA: "L2_metadata/"
  logfile: ""
date: now
date-format: "YYYY-MM-DD HH:mm:ssZ"
format: 
  html:
    embed-resources: true
    code-fold: true
editor: visual
---

This script

-   Reads in the L1 data files one by one

-   Drops NA data rows (shouldn't be any?)

-   Drops flagged data: `F_OOB`, `F_OOS`, outliers `F_MAD`

-   Averages to `r params$timestamp_round`, logging N_avg and N_drop

-   Completes data for entire year, ensuring continuous time series

-   Interpolates short data gaps

-   Writes L2_qaqc files

## Initializing

```{r init}
#| include: false

library(tidyr)
library(readr)
library(lubridate)
library(zoo)
library(ggplot2)

L1 <- file.path(params$DATA_ROOT, params$L1)
L2_QAQC <- file.path(params$DATA_ROOT, params$L2_QAQC)

files_to_process <- list.files(L1, 
                               pattern = "[A-Z]{3}_[A-Z]+[1-3]*_[0-9]{8}-[0-9]{8}_.+csv$", 
                               full.names = TRUE, recursive = TRUE)

QAQC_TABLE <- file.path(params$METADATA_ROOT, params$QAQC_TABLE)
qaqct <- read_csv(QAQC_TABLE, col_types = "cccccc")
qaqct <- qaqct[!is.na(qaqct$variable),]

# We use the data period (frequency) in gap filling below
data_period <- period(params$timestamp_round)
if(is.na(data_period)) {
    stop("Invalid timestamp_round value: ", params$timestamp_round)
}

source("helpers.R")
source("L2_qaqc-utils.R")
```

I see `r length(files_to_process)` files to process in `r L1`.

QAQC table `r QAQC_TABLE` and has `r nrow(qaqct)` entries.

Round data to `r params$timestamp_round`.

Output directory is `r L2_QAQC`

HTML outfile is "`r params$html_outfile`".

## Processing

```{r processing}
overwrites <- 0
errors <- 0
dat_issues <- 0

f <- function(fn, out_dir) {
    message(Sys.time(), " Processing ", basename(fn))

    dat <- read_csv(fn, col_types = "ccTccccccciii")
    rn <- dat$research_name[1]
    message("\t", rn, " total data: ", nrow(dat), " rows, ", ncol(dat), " columns")
    
    # File-based summary
    smry <- data.frame(File = basename(fn), 
                       Site = dat$Site[1],
                       Plot = dat$Plot[1],
                       research_name = rn,
                       Year = unique(year(dat$TIMESTAMP)),
                       Rows = nrow(dat))
    
    # loop through all the F_ flags, count, and add to summary table,
    flag_cols <- grep("^F_", colnames(dat))
    dat$F_drop <- rep(0L, nrow(dat)) # signals whether to drop row or not
    for(fname in colnames(dat)[flag_cols]) {
        # NA flags are treated as false (zero)
        dat[[fname]][is.na(dat[[fname]])] <- FALSE
        
        f_smry <- data.frame(sum(dat[[fname]], na.rm = TRUE))
        colnames(f_smry) <- fname
        smry <- cbind(smry, f_smry)
        
        dat$F_drop <- dat$F_drop + dat[[fname]]
    }
    smry$Total <- sum(dat$F_drop)
    smry$Pct <- paste0(sprintf("%6.1f", smry$Total / nrow(dat) * 100), "%")
    
    # Round timestamps and average
    dat$TIMESTAMP <- round_date(dat$TIMESTAMP, unit = params$timestamp_round)
    dat_summarised <- L2_aggregate(dat)
    
    # Complete the data: fill in (with NA) all timestamps and all combinations
    # of Instrument, Sensor, etc. This guarantees a smooth time series for
    # each sensor
    rows <- nrow(dat_summarised)
    dat_complete <- L2_complete(dat_summarised) # function is in L2_qaqc-utils
    message("\tCompleted data; started with ", rows, 
            " rows...now ", nrow(dat_complete))
    
    # Interpolate across small gaps for each site, plot, instrument, and sensor
        
    # 'maxgap' controls zoo::na_approx's gap-filling behavior; we compute it
    # based on the data interval and the max_gap_fill entry in the QAQC table
    if(is.numeric(dat_complete$Value)) {
        i <- which(qaqct$variable == rn)
        od_period <- period(qaqct$max_interp[i])
        if(is.na(od_period)) {
            stop("Invalid max_gap_fill entry for ", rn)
        }
        maxgap <- od_period / data_period # e.g., 1 hour / 15 min = 4
        na_vals <- is.na(dat_complete$Value)
        message("\tGap-filling data with maxgap = ", maxgap)
        dat_interped <- L2_linear_interp(dat_complete, maxgap = maxgap)
        new_na_vals <- is.na(dat_interped$Value)
        message("\t\tStarted with ", sum(na_vals), 
                " missing values...now ", sum(new_na_vals))
        smry$Filled <- sum(na_vals & !new_na_vals)
    } else {
        message("\tNo gap-filling; Value is not numeric")
        dat_interped <- dat_complete
        smry$Filled <- 0L
    }
    
    message("\tWriting...")
    write_to_folders(dat_interped,
                     out_dir,
                     "L2_qaqc",
                     site = dat_interped$Site[1],
                     plot = dat_interped$Plot[1],
                     write_plots = FALSE) 
    
    return(smry)
}

log_info("About to L2_qaqc", logfile = params$logfile)
tryCatch({
    out <- lapply(files_to_process, f, out_dir = L2_QAQC)
},
error = function(e) {
    log_warning("L2: an error occurred!", logfile = params$logfile)
    log_info(as.character(e), logfile = params$logfile)
    stop(e)
})
```

## Summary

```{r summary}
#| echo: false
#| output: asis
if(errors) {
    cat("### WARNING: ", errors, " file read/write error(s)\n")
    log_warning(paste("File read/write error(s)", params$html_outfile), 
                logfile = params$logfile)
}
```

```{r summary_table}
out_df <- do.call("rbind", out)
knitr::kable(out_df[-1]) # don't show filename
```

## Reproducibility

Git commit `r GIT_COMMIT`.

```{r reproducibility}
sessionInfo()
```
